Gender and international diversity improves equity in peer review: R Notebook

### 2018-08-06 eLife_Analysis_and_Figures.Rmd
Updated R Notebook

---
title: 'Gender and international diversity improves equity in peer review: R Notebook'
author: "Dakota Murray"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

@@

@@

### Final decision outcomes for full submissions
```{r}
outcomes <- full %>%
  mutate(outcome = ifelse(!final_decision_made, "No Decision Yet", ifelse(accepted, "Accepted", "Rejected")))

@@

### First full decision outcomes for full submissions
```{r}
table(ifelse(is.na(full$full_decision), "No Decision", full$full_decision))
```

### First revision review outcomes for full submissions
```{r}
rev1 <- full %>% filter(!is.na(rev1_submission_date))
table(ifelse(is.na(rev1$rev1_decision), "No Decision", rev1$rev1_decision))
```

### Second revision review outcomes for full submissions
```{r}
rev2 <- full %>% filter(!is.na(rev2_submission_date))
table(ifelse(is.na(rev2$rev2_decision), "No Decision", rev2$rev2_decision))
```

### Number of distinct corresponding authors in full submissions at eLife
```{r}
ca_authors <- full %>% 
  filter(!is.na(ca_name)) %>%
  mutate(ca_name = tolower(ca_name)) %>%
  group_by(ca_name, ca_country) %>%
  filter(row_number() == 1)

@@

### Gender distirbution of distinct correpsonding authors on full submissions at eLife
```{r}
table(ca_authors$ca_gender)
```

@@

### Gender distirbution of distinct first authors on full submissions at eLife
```{r}
table(fa_authors$fa_gender)
```

### Number of distinct last authors on full submissions at eLife
```{r}
la_authors <- full %>% 
  filter(!is.na(la_name)) %>%
  mutate(la_name = tolower(la_name)) %>%
  group_by(la_name, la_country) %>%
  filter(row_number() == 1)

@@

@@

### Gender distirbution of distinct gatekeepers involved in processing full submissions at eLife, includes senior editors, reviewing editors, and invited peer reviewers
```{r}
table(gatekeepers$gender)
```

### Gender distirbution of distinct gatekeepers involved in processing full submissions at eLife, includes senior editors, reviewing editors, and invited peer reviewers as %
```{r}
prop.table(table(gatekeepers$gender))
```

### Continent distirbution of distinct gatekeepers involved in processing full submissions at eLife, includes senior editors, reviewing editors, and invited peer reviewers
```{r}
table(gatekeepers$continent)
```

### Proportion of full submissions for which the correposnding author was also the last author
```{r}
sum(full$ca_name == full$fa_name, na.rm = T) / dim(full)[1]
```

@@

@@

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
gatekeepers <- subset(people, type == "Gatekeeper")
#table(gatekeepers$continent)
editors <- subset(editors, !duplicated(name, country))
reviewers <- subset(reviewers, !duplicated(reviewer_name, reviewer_country))
bres <- subset(bres, !duplicated(name, country))

@@

@@

And the same for the content-level tests
```{r}
fig3_bottom_sig_table <- data.frame(
  continent = c("Corresponding", "First", "Last"),
  p.value = sapply(continent_sig_list, function(x) x$p.value),
  n = sapply(continent_sig_list, function(x) sum(x$observed)),
  statistic = sapply(continent_sig_list, function(x) x$statistic),
  df = sapply(continent_sig_list, function(x) x$parameter)
)

@@

### Presence of author-reviewer national homogeny by country of correposnding author
```{r rows.print = 20}
full <- elife %>% 
  filter(has_full_submission) %>%
  group_by(ca_country) %>%
  summarize(
    submissions = n(),
    prop_homogeny = sum(ca_country_homophily, na.rm = T) / submissions
  ) %>%
  arrange(-prop_homogeny)

full
```

>>>>

---
title: 'Gender and international diversity improves equity in peer review: R Notebook'
author: "Dakota Murray"
output:
  pdf_document: default
  html_notebook: default
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
  html_document:
    df_print: paged
---

@@

## Proportion of initial submissions with decisions which were accepted/rejected
```{r}
with_decision <- outcomes %>% filter(outcome != "No Decision Yet")

prop.table(table(with_decision$outcome))
```

@@

###Final decision outcomes for full submissions
```{r}
outcomes <- full %>%
  mutate(outcome = ifelse(!final_decision_made, "No Decision Yet", ifelse(accepted, "Accepted", "Rejected")))

@@

### First full decision outcomes for full submissions
```{r}
table(ifelse(is.na(full$full_decision), "No Decision", full$full_decision))
prop.table(table(ifelse(is.na(full$full_decision), "No Decision", full$full_decision)))
```

###First revision review outcomes for full submissions
```{r}
rev1 <- full %>% filter(!is.na(rev1_submission_date))
table(ifelse(is.na(rev1$rev1_decision), "No Decision", rev1$rev1_decision))
```

###Second revision review outcomes for full submissions
```{r}
rev2 <- full %>% filter(!is.na(rev2_submission_date))
table(ifelse(is.na(rev2$rev2_decision), "No Decision", rev2$rev2_decision))
```

###Number of distinct corresponding authors in full submissions at eLife
```{r}
ca_authors <- full %>% 
  filter(!is.na(ca_name)) %>%
  mutate(ca_name = tolower(ca_name)) %>%
  group_by(ca_name, ca_country) %>%
  filter(row_number() == 1)

@@

###Gender distirbution of distinct correpsonding authors on full submissions at eLife
```{r}
table(ca_authors$ca_gender)
```

@@

### Gender distirbution of distinct first authors on full submissions at eLife
```{r}
table(fa_authors$fa_gender)
prop.table(table(fa_authors$fa_gender))
```

###Gender distirbution of distinct gender-identified first authors on full submissions at eLife
```{r}
identified <- fa_authors %>% filter(fa_gender %in% c("M", "F"))
table(identified$fa_gender)
prop.table(table(identified$fa_gender))
```

###Number of distinct last authors on full submissions at eLife
```{r}
la_authors <- full %>% 
  filter(!is.na(la_name)) %>%
  mutate(la_name = tolower(la_name)) %>%
  group_by(la_name, la_country) %>%
  filter(row_number() == 1)

@@

### Gender distirbution of distinct gender-identified first authors on full submissions at eLife
```{r}
identified <- la_authors %>% filter(la_gender %in% c("M", "F"))
table(identified$la_gender)
prop.table(table(identified$la_gender))
```

@@

### Gender distirbution of distinct gatekeepers involved in processing full submissions at eLife, includes senior editors, reviewing editors, and invited peer reviewers
```{r}
table(gatekeepers$gender)
prop.table(table(gatekeepers$gender))
```

###Gender distribution of distinct gatekeepers invovled in the processing of full submissions submissions at eLife, but now for only gender-identified individuals
```{r}
identified <- gatekeepers %>% filter(gender %in% c("M", "F"))
table(identified$gender)
prop.table(table(identified$gender))
```

### Continent distirbution of distinct gatekeepers involved in processing full submissions at eLife, includes senior editors, reviewing editors, and invited peer reviewers
```{r}
table(gatekeepers$continent)
prop.table(table(gatekeepers$continent))
```

### Proportion of full submissions for which the correposnding author was also the first author
```{r}
sum(full$ca_name == full$fa_name, na.rm = T) / dim(full)[1]
```

@@

###Number of unique name/role combinations that appear in the data
```{r}
dim(people)[1]
```

###Number of thees combinations for which a gender was assigned
```{r}
sum(people$gender != "UNK")
sum(people$gender != "UNK") / dim(people)[1]

```

###Gender distribution of all name/role combinations that appear in the dataset
```{r}
table(people$gender)
prop.table(table(people$gender))
```

@@

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
gatekeepers <- subset(people, type == "Gatekeeper")
editors <- subset(editors, !duplicated(name, country))
reviewers <- subset(reviewers, !duplicated(reviewer_name, reviewer_country))
bres <- subset(bres, !duplicated(name, country))

@@

###Gender distribution of gender-identified correpsonding authors of full submissions
```{r}
identified <- full %>% filter(ca_gender != "UNK")
table(identified$ca_gender)
prop.table(table(identified$ca_gender))
```

###Gender distribution of gender-identified first authors of full submissions
```{r}
identified <- full %>% filter(fa_gender != "UNK")
table(identified$fa_gender)
prop.table(table(identified$fa_gender))
```

###Gender distribution of gender-identified last authors of full submissions
```{r}
identified <- full %>% filter(la_gender != "UNK")
table(identified$la_gender)
prop.table(table(identified$la_gender))
```

@@

And the same for the continent-level tests
```{r}
fig3_bottom_sig_table <- data.frame(
  continent = c("Corresponding", "First", "Last"),
  p.value = sapply(continent_sig_list, function(x) x$p.value),
  n = sapply(continent_sig_list, function(x) sum(x$observed)),
  statistic = sapply(continent_sig_list, function(x) x$statistic),
  df = sapply(continent_sig_list, function(x) x$parameter)
)

@@

### Presence of author-reviewer national homogeny by country of correposnding author
```{r rows.print = 20}
full <- elife %>% 
  filter(has_full_submission) %>%
  group_by(ca_country) %>%
  summarize(
    submissions = n(),
    prop_homogeny = sum(ca_country_homophily, na.rm = T) / submissions
  ) %>%
  arrange(-prop_homogeny) %>%
  top_n(20)

full

write.csv(full, file = "tables/table_SI3.csv")
```
